Modeling memory and taxis to analyze migrations

### 2021-05-14 text/ms.Rmd
working on ms ...

Migration is a strategic behavior, typically involving a major trade-off between the costs of the migration against the benefits of accessing resources (whether food or predator refuge) that are highly seasonal and localized (**refs**).  While some migrants - both terrestrial and avian - follow a "green wave" of spring vegetation as it flowers across altitudinal or latitudinal gradients [@Bischof2012;@Kolzsch2015], many others "jump" the green wave, anticipating that critical resources will be available at the time of arrival (**refs**).

We anticiapted that under many conditons a blending of *tactical* (i.e. direct response to resource availability or perception) and *strategic* (i.e. memory-driven and forward-thinking) behavior can help a forager navigate a dynamic and heterogeneous resource environment, with a particular emphasis on the aquisition and adaptabilty of **migratory behavior** in a dynamic, seasonal resource environment To explore the interaction of memory and resource following as an adaptive strategy in dynamic environments. Specifically, we explore a one-dimensional advection diffusion model with long-term and short-term memory, and run the model in a dynamic, seasonal environment.

>>>>

Seasonal migrations are an extremely widespread among terrestrial, aquatic, avian and invertebrate species [@Dingle2014]. For many species, it is an extremely successful strategy as it allows a greater number of individuals to inhabit landscape which might not otherwise be able to support large numbers year round [@Fryxell1996]. The proximate and ultimate causes and mechanisms, however, vary considerably across species, and even within species [@Shaw2016].Some migrants follow a "green wave" of spring vegetation as it flowers across altitudinal or latitudinal gradients [@Bischof2012;@Kolzsch2015] and can be considered "tactical" migrants.  Others "jump" a green wave (or any other resource wave) anticipating that critical resources will be available at the time of arrival at the end point of migration (**refs**).  This second behavior involves a major trade-off between the costs of migration against the benefits of accessing resources, whether food, suitable habitat for breeding, or predator refuge, that are highly seasonal and localized (**refs**).  In it's can be considered "strategic" in the sense that it is driven not by immediate cues but by memory-driven anticipation.  Clearly,

While migratory species are often considered to be more vulnerable to environmental change, as a disruptions in either of the seasonal ranges or along a migratory corridor can have significant negative impacts [@Wilcove2008;@Kauffman2021].  On the other hand, it has been argued that migratory species might be more resilient to dispruptions due to their wide-ranging mobility [@Robinson2009].

We anticiapte that under many conditons a blending of *tactical* (i.e. direct response to resource availability or perception) and *strategic* (i.e. memory-driven and forward-thinking) behavior can help a forager navigate a dynamic and heterogeneous resource environment, with a particular emphasis on the aquisition and adaptabilty of *migratory behavior* in a dynamic, seasonal resource environment. Our goal is to explore the interaction of memory and resource following as an adaptive strategy in a dynamic environments. To do this, we developed a one-dimensional advection diffusion model with long-term and short-term memory, and run the model in a dynamic, seasonal environment.
